TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-07 paper/draft.tex
describe previous m dwarf studies

\apogee\ is primarily targeted for bright stellar populations particularly red giants, with de-reddened photometry and color cutoffs of $7 \leq H \leq 13.8$ and $[J-K]_0 \geq 0.5$ \citep{Zasowski:2013} with the objective of studying galactic composition and evolution. However numerous cool, main sequence sources have also been observed either as targets proposed by M dwarf ancilliary surveys, or serendipitously. We determine in this study at least 10,000 M dwarfs have been observed by \apogee\ to date.

@@

\color{gcolor}{JB: I've moved this paragraph from the intro to here. Will expand this... A number of M dwarf studies have already been conducted in \apogee, particularly towards measuring reliable fundamental atmospheric parameters and kinematic measurements. \citealt{Schmidt:2016} -- analyzed trends and reliablility of ASPCAP parameters for K/M dwarfs. \citealt{Souto:2017} and \citealt{Souto:2018} -- modeled three exoplanet hosting M dwarfs (Kepler-138, Kepler-168 and Ross-128), determining Teff/logg/metallicity + 13 elemental abundances. \citealt{Desphande:2013} -- modeled rotational and radial velocities for 200+ M dwarfs, later extended by \citealt{Gilhool:2018} to 700+ M dwarfs. \citealt{Rajpurohit:2018} -- tested BTSettl model grids on 45 M dwarfs, estimated Teff/logg/metallicity). \citealt{Skinner:2018} - identified and measured mass ratios and radial velocities for 44 M dwarf binaries.}\color{black}

>>>>

\apogee\ is primarily targeted for bright stellar populations particularly red giants, with de-reddened photometry and color cutoffs of $7 \leq H \leq 13.8$ and $[J-K]_0 \geq 0.5$ \citep{Zasowski:2013} with the objective of studying galactic composition and evolution. However numerous cool, main sequence sources have also been observed either as targets proposed by the \apogee\ M dwarf ancilliary survey ($\sim$1,200 sources; \citealt{Desphande:2013}), or serendipitously.

A number of studies out of the M dwarf ancilliary survey have already been conducted towards measuring reliable fundamental atmospheric parameters and kinematic measurements using spectral synthesis of atmospheric model grids. 
These studies include \citealt{Desphande:2013} and \citealt{Gilhool:2018} which have studied the radial and rotational kinematics for 700+ sources,
\citealt{Souto:2017} and \citealt{Souto:2018} which have modeled three exoplanet hosting M dwarfs (Kepler-138, Kepler-168 and Ross-128), determining $\teff$/$\logg$/metallicity + 13 elemental abundances,
\citealt{Rajpurohit:2018} which tested BTSettl model grids \citep{Allard:2012} on 45 M dwarfs to estimate $\teff$/$\logg$/metallicity,
and \citealt{Skinner:2018} which has identified and measured mass ratios and radial velocities for 44 M dwarf spectroscopic binaries.
This work complements existing studies by producing model-independent catalog of spectroscopic temperatures and metallicites to test against model predictions for the entire \apogee\ M dwarf sample, which we quantify to contain at least 10,000 sources to date (DR14).

@@
